---
title: "R6"
author: "Zach Schuster"
date: "4/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(R6)
```

## 14.2.6

1. Create a bank account R6 class that stores a balance and allows you to deposit and withdraw money. Create a subclass that throws an error if you attempt to go into overdraft. Create another subclass that allows you to go into overdraft, but charges you a fee.

```{r}

# bank account class
Account = R6Class(
  "Account",
  public = list(
    balance = NULL,
    
    initialize = function(balance) {
      self$balance = balance
      invisible(self)
    },
    deposit = function(amount) {
      self$balance = self$balance + amount
      invisible(self)
    },
    withdraw = function(amount) {
      self$balance = self$balance - amount
      invisible(self)
    }
  ))

zachs_account = Account$new(balance = 100)
zachs_account$withdraw(50)
zachs_account
zachs_account$deposit(100)
zachs_account

# This class doesn't allow you to go into overdraft
AccountOverdraft = R6Class(
  "AccountOverdraft",
  inherit = Account,
  public = list(
    withdraw = function(amount) {
      new_balance = self$balance - amount
      if (new_balance < 0) {
        stop("You can't go into overdraft. Take out less money!")
      } else {
        super$withdraw(amount)
      }
    }
  )
)

ovdft = AccountOverdraft$new(100)
ovdft$withdraw(101)
ovdft$withdraw(99)
ovdft


# You can go into overdraft with a fee. This assumes the balance before withdrawal is above 0.
AccountOverdraftFee = R6Class(
  "AccountOverdraftFee",
  inherit = Account,
  public = list(
    withdraw = function(amount) {
      new_balance = self$balance - amount
      if (new_balance < 0) {
        amount_over = 0 - new_balance
        fee = 0.03 * amount_over
        self$balance = new_balance - fee
        print("You've been charged a 3% fee on overdrafted funds.")
        invisible(self)
      } else {
        super$withdraw(amount)
      }
    }
  )
)

ovdft = AccountOverdraftFee$new(100)
ovdft$withdraw(101)
ovdft
ovdft = AccountOverdraftFee$new(100)
ovdft$withdraw(99)
ovdft

```

2. Create an R6 class that represents a shuffled deck of cards. You should be able to draw cards from the deck with $draw(n), and return all cards to the deck and reshuffle with $reshuffle(). Use the following code to make a vector of cards.

```{r}

CardDeck = R6Class(
  "CardDeck",
  public = list(
    # defined by hadley
    suit = c("♠", "♥", "♦", "♣"),
    value = c("A", 2:10, "J", "Q", "K"),
    cards = paste0(rep(value, 4), suit),
    
    draw = function(n) {
      pulled = sample(cards, size = n, replace = FALSE)
      self$cards = setdiff(self$cards, pulled)
      return(pulled)
    },
    reshuffle = function() {
      # Doing this in two steps for readability. recreate full deck.
      self$cards = paste0(rep(self$value, 4), self$suit)
      # reshuffle the cards
      self$cards = sample(self$cards)
    }

  )
)

zachs_deck = CardDeck$new()
zachs_deck$draw(5)
length(zachs_deck$cards)
zachs_deck$reshuffle()
length(zachs_deck$cards)
```

3. Why can’t you model a bank account or a deck of cards with an S3 class?

A few reasons. 

* The balance of a bank account is all tied to the instance of the class, which makes it much easier to keep track of deposits and withrawals and balances. It is all contained.
* Having reference semantics in R6 makes life way more cenvenient and efficient.

4. Create an R6 class that allows you to get and set the current timezone. You can access the current timezone with Sys.timezone() and set it with Sys.setenv(TZ = "newtimezone"). When setting the time zone, make sure the new time zone is in the list provided by OlsonNames().

```{r}

Timezone = R6Class(
  "Timezone",
  public = list(
    # get all time zones that can be set
    time_zones = OlsonNames(),
    # This sets the original time zone as EST for our purposes.
    original_zone = "America/New_York",
    get_tz = function() {
      return(Sys.timezone())
    },
    set_tz = function(zone = "America/New_York") {
      Sys.setenv(TZ = zone)
      # This invisibly returns TRUE
    },
    reset_tz = function() {
      Sys.setenv(TZ = self$original_zone)
    }
  ))

zone = Timezone$new()

zone$time_zones

zone$get_tz()
zone$set_tz("Poland")
Sys.time()
zone$get_tz()

zone$reset_tz()
Sys.timezone()

```

5. Create an R6 class that manages the current working directory. It should have $get() and $set() methods.

```{r}

WDManager = R6Class(
  "WDManager",
  public = list(
    get = function() {
      return(getwd())
    },
    set = function(path = "~") {
      setwd(path)
    }
  )
)

# store the old working directory
old_wd = getwd()
wd_manager = WDManager$new()

wd_manager$get()

wd_manager$set()
wd_manager$get()

# restore working directory
wd_manager$set(old_wd)

# double check with base R
getwd()
```

6. Why can’t you model the time zone or current working directory with an S3 class?

It is not as easy to use functionality like get or set. It would then become get.tz or something similar so there has to be an underlying class that would be input to the function. Not nearly as simple in my opinion. 

7. What base type are R6 objects built on top of? What attributes do they have?

```{r}

TestClass = R6Class(
  "TestClass",
  public = list()
)

class(unclass(TestClass))
class(unclass(TestClass$new()))

# R6 is build on an environment

attributes(TestClass)
attributes(TestClass$new())

```






















