---
title: "S3 System"
author: "Zach Schuster"
date: "3/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sloop)
```

## Intro

"I’d recommend coupling the theoretical knowledge from this chapter with the practical knowledge encoded in the vctrs package."

* The implementation for a specific class is called a method, and the generic finds that method by performing method dispatch.

## 12.2.1

1. Describe the difference between t.test() and t.data.frame(). When is each function called?

`t.test` is called as the disbatch function for any subsequent methods. It is a generic function. It is not a method, shown below. `t.data.frame` is the transpose method for class data.frame.

```{r}
ftype(t.test)
ftype(t.data.frame)

is_s3_method("t.test")
is_s3_method("t.data.frame")
```

2. Make a list of commonly used base R functions that contain . in their name but are not S3 methods.

```{r}
is_s3_method("weighted.mean")
is_s3_method("data.frame")
is_s3_method("prop.table")
is_s3_method("binom.test")
# other suggestions?
```

3. What does the `as.data.frame.data.frame()` method do? Why is it confusing? How could you avoid this confusion in your own code?

It converts a data.frame to a data.frame!It is really confusing because of all the periods. I'd use underscores in the function name to separate methods (seems only logical). To avoid it in code, just call the generic function and let method disbatch work behind the scenes. 

```{r}
df = data.frame(a = 1:5,
                b = 6:10)

as.data.frame.data.frame(df)
identical(df, as.data.frame.data.frame(df))
```

4. Describe the difference in behaviour in these two calls.

```{r}
set.seed(1014)
some_days <- as.Date("2017-01-31") + sample(10, 5)

mean(some_days)
#> [1] "2017-02-05"
mean(unclass(some_days))
#> [1] 17202
```


The first call is being disbatched to the `mean.Date` method whereas the second call is being disbatched to `mean.default` because there is no specific method for class numeric, so it goes to `mean.default`.

5. What class of object does the following code return? What base type is it built on? What attributes does it use?

```{r}
x <- ecdf(rpois(100, 10))
x

# class
class(x)

# base type # closure
typeof(x)

# also considered a function is R terms # function
class(unclass(x))

# what attributes does it use? # class, call
attributes(x)
```

6. What class of object does the following code return? What base type is it built on? What attributes does it use?

```{r}
x <- table(rpois(100, 5))
x

class(x) # table
class(unclass(x)) # underlying is an array
```


## 13.3.4

1. Write a constructor for data.frame objects. What base type is a data frame built on? What attributes does it use? What are the restrictions placed on the individual elements? What about the names?

* data.frame is built on a list
* it uses the class, names, and row.names attributes
* all elements of the list must be the same length. Every element of list must only be one class
* names can not have special characters

```{r}
new_data.frame = function(x = list(),
                          col_names = character(), # names attr
                          row_names = character()) { # row.names attr
  
  stopifnot(is.character(col_names))
  stopifnot(is.character(row_names))
  
  structure(
    x,
    class = 'data.frame',
    names = col_names,
    row.names = row_names
  )
}

(new_df = new_data.frame(list(1:5, 6:10),
               col_names = c("a", "b"),
               row_names = as.character(1:5))
)
class(new_df)
```

2. Enhance my factor() helper to have better behaviour when one or more values is not found in levels. What does base::factor() do in this situation?

```{r}
validate_factor <- function(x) {
  values <- unclass(x)
  levels <- attr(x, "levels")

  if (!all(!is.na(values) & values > 0)) {
    stop(
      "All `x` values must be non-missing and greater than zero",
      call. = FALSE
    )
  }

  if (length(levels) < max(values)) {
    stop(
      "There must at least as many `levels` as possible values in `x`",
      call. = FALSE
    )
  }

  x
}


factor <- function(x = character(), levels = unique(x)) {
  ind <- match(x, levels)
  validate_factor(new_factor(ind, levels))
}
```


3. Carefully read the source code of factor(). What does it do that my constructor does not?

```{r}
factor
```

* There appears to be more in depth error checking
* There is the option to make an ordered factor
* The error messages are generalized

4. Factors have an optional “contrasts” attribute. Read the help for C(), and briefly describe the purpose of the attribute. What type should it have? Rewrite the new_factor() constructor to include this attribute.

for our purposes, it should be a matrix. It allows us to compare different groups of levels within a factor, usually done through a lm or glm or aov.

```{r}
new_factor <- function(x = integer(), levels = character(),
                       contrasts = NULL) {
  stopifnot(is.integer(x))
  stopifnot(is.character(levels))

  fac = structure(x,
                  levels = levels,
                  class = "factor")
  
  # add contrasts if there are any
  if (!is.null(contrasts && is.matrix(contrasts))) {
    if (length(levels) == nrow(contrasts)) {
      fac = C(fac, contr = contrasts)
      return(fac)
    }
  } else {
      stop("contrasts should be a matrix with nrow == number of levels")
    }
}

# make contrasts
cont = matrix(c(1, 1, -1, -1,
                1, 1, -1, -1,
                -1, -1, 1, 1,
                -1, -1, 1, 1),
              byrow = TRUE, nrow = 4)

(x = new_factor(rep(1L:4L, each = 3),
           levels = c('apple', 'orange', 'grean_beans', 'broccoli'),
           contrasts = cont))
```


5. Read the documentation for utils::as.roman(). How would you write a constructor for this class? Does it need a validator? What might a helper do?

```{r}
?utils::as.roman
```

You could write a constructor function that take an integer or character vector as an argument and use the str class to assign the class 'roman' to it. 

A validator would be nice to make sure numbers are between 1 and 3899

A helper could help with converting numeric to integer or something like that. 


## 13.4.4

1. Read the source code for t() and t.test() and confirm that t.test() is an S3 generic and not an S3 method. What happens if you create an object with class test and call t() with it? Why?

```{r}
t
t.test
sloop::is_s3_generic('t.test')

x = structure(1:10, class = 'test')
t(x)

sloop::s3_dispatch(t(x))
sloop::s3_dispatch(t.test(x))
```

Because `t` also is a generic, it searches through possible methods. `t.test` is a possible option, so it gets called. I believe `t.test` then calls method dispatch, which calls the `t.test.default`. 

2. What generics does the table class have methods for?

```{r}
s3_methods_class("table")
```

3. What generics does the ecdf class have methods for?

```{r}
s3_methods_class("ecdf")
```

4. Which base generic has the greatest number of defined methods?

```{r}
builtin = builtins()
ind = unlist(lapply(builtin, function(x) is.function(get(x))))
base_funs = builtin[ind]

# get base generics
base_generics = base_funs[vapply(base_funs, is_s3_generic, FUN.VALUE = logical(1))]

# how many methods does each have?
num_args = unlist(lapply(base_generics, function(x) nrow(s3_methods_generic(x))))
base_generics[which.max(num_args)]

s3_methods_generic('print')
```

5. 







