---
title: "Advanced R: Rcpp"
author: "Zach Schuster"
date: "9/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(microbenchmark)
library(Rcpp)
```

```{r}

cppFunction(
  'int add(int x, int y, int z) {
    int sum = x + y + z;
    return sum;
  }'
)

add_r = function(x, y, z) {
  x + y + z
}

microbenchmark(
  add(1, 4, 7),
  add_r(1, 4, 7),
  times = 1e6
)

```

here R is faster


### Function Comparison

#### no input, scalar output

```{r}

# R
one = function() 1L

# C++
cppFunction('
int one() {
  return 1;
}'
)

```


* you don't use assignment to create a funtion in C++  
* explicit return statement is necessary
* you must use `;` after every statement in C++


#### scalar input, scalar output

```{r}

# R
signR = function(x) {
  if (x > 0) {
    1
  } else if (x == 0) {
    0
  } else {
    -1
  }
}

# C++
cppFunction('
            int signC(int x) {
              if (x > 0) {
                return 1;
              } else if (x == 0) {
                return 0;
              } else {
                return -1;
              }
            }
')

microbenchmark(
  signR(5),
  signC(5),
  times = 1e5
)

```

again we see here that R is much faster than C++


#### vector input, scalar output

```{r}

# R. for comparison, not how sum should actually be implemented. I've included a
# for loop implementation that no longer uses subsetting as well
sumR = function(x) {
  total = 0
  for (i in seq_along(x)) {
    total = total + x[i]
  }
  total
}

sum_quick_R = function(x) {
  total = 0
  for (num in x) {
    total = total + num
  }
  total
}

cppFunction(
  'double sumC(NumericVector x) {
    int n = x.size();
    double total = 0;
    for (int i = 0; i < n; ++i) {
      total += x[i];
    }
    return total;
  }'
)

```


* .size() method finds the length of the vector

* The for statement has a different syntax: for(init; check; increment). This loop is initialised by creating a new variable called i with value 0. Before each iteration we check that i < n, and terminate the loop if it’s not. After each iteration, we increment the value of i by one, using the special prefix operator ++ which increases the value of i by 1.


```{r}

x = runif(1e5)
microbenchmark(
  sumR(x),
  sum_quick_R(x),
  sum(x),
  sumC(x)
)

```

The c++ function is about 25-40x faster than the equivalent R function. For loops appear to be a place where C++ can come in handy


#### Vector Input, Vector Output

* create a function that takes in a scalar and a vector and calculates euclidean distance between scalar and each element of the vector

```{r}
# function input types are not clear in R
pdistR = function(x, ys) {
  (sqrt(x - ys) ^ 2)
}

cppFunction('NumericVector pdistC(double x, NumericVector ys) {
              int n = ys.size();
              NumericVector out(n);
              
              for (int i = 0; i < n; ++i) {
                out[i] = sqrt(pow(ys[i] - x, 2));
              }
              return out;
            }
')

x = 7.5
ys = rnorm(1e6)

microbenchmark(
  pdistR(x, ys),
  pdistC(x = x, ys = ys)
)
```

* Note that because the R version is fully vectorised, it’s already going to be fast. On my computer, it takes around 8 ms with a 1 million element y vector. The C++ function is twice as fast, ~4 ms, **but assuming it took you 10 minutes to write the C++ function, you’d need to run it ~150,000 times to make rewriting worthwhile**. The reason why the C++ function is faster is subtle, and relates to memory management. The R version needs to create an intermediate vector the same length as y (x - ys), and allocating memory is an expensive operation. The C++ function avoids this overhead because it uses an intermediate scalar.


#### Matrix Input, Vector Output

reproduce rowSums

```{r}
cppFunction('NumericVector rowSumsC(NumericMatrix x) {
  int nrow = x.nrow(), ncol = x.ncol();
  NumericVector out(nrow);
  
  for (int i = 0; i < nrow; ++i) {
    for (int j = 0; j < ncol; ++j) {
      out[i] += x(i, j);
    }
  }
  return out;
}
')

x = matrix(rnorm(1e4), nrow = 10)

microbenchmark(
  rowSums(x),
  rowSumsC(x)
)

```

`rowSumsC` is about twice as fast


## sourceCpp

* must start with:

`#include <Rcpp.h>`

`using namespace Rcpp;`

```{r}

sourceCpp('mean.cpp')

```

### Exercises 19.1.7

What R function does the c++ code represent?

```{cpp eval = FALSE}

double f1(NumericVector x) {
  int n = x.size();
  double y = 0;

  for(int i = 0; i < n; ++i) {
    y += x[i] / n;
  }
  return y;
}

// mean

NumericVector f2(NumericVector x) {
  int n = x.size();
  NumericVector out(n);

  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
    out[i] = out[i - 1] + x[i];
  }
  return out;
}

// cumsum

bool f3(LogicalVector x) {
  int n = x.size();

  for(int i = 0; i < n; ++i) {
    if (x[i]) return true;
  }
  return false;
}

// any

int f4(Function pred, List x) {
  int n = x.size();

  for(int i = 0; i < n; ++i) {
    LogicalVector res = pred(x[i]);
    if (res[0]) return i + 1;
  }
  return 0;
}

// Position (had originally guessed which)

NumericVector f5(NumericVector x, NumericVector y) {
  int n = std::max(x.size(), y.size());
  NumericVector x1 = rep_len(x, n);
  NumericVector y1 = rep_len(y, n);

  NumericVector out(n);

  for (int i = 0; i < n; ++i) {
    out[i] = std::min(x1[i], y1[i]);
  }

  return out;
}

// pmin

```



practice writing functions in cpp

1. all()
2. cumprod(), cummin(), cummax()
3. diff()
4. range()
5. var()

__refer to exercise_funcs.cpp for source code__

```{r}
Rcpp::sourceCpp('exercise_funcs.cpp')
```


### Attributes and other classes

Look at `attribs.cpp` for example

```{r}
Rcpp::sourceCpp('attribs.cpp')
```


### Missing Values

#### scalars

```{r}
cppFunction(
  "List scalar_missings() {
  int int_s = NA_INTEGER;
  String chr_s = NA_STRING;
  bool lgl_s = NA_LOGICAL;
  double num_s = NA_REAL;
  
  return List::create(int_s, chr_s, lgl_s, num_s);
  }"
)

str(scalar_missings())
```

#### Show in online book some of the finicky examples with missing values. 

### 19.3.5 Exercises

 Only did Q1

```{r}
Rcpp::sourceCpp('var_narm.cpp')
```


## Rcpp Sugar

__If there’s a sugar version of the function you’re interested in, you should use it: it’ll be both expressive and well tested.__
 
Sugar functions can be roughly broken down into

* arithmetic and logical operators
* logical summary functions
* vector views
* other useful functions

```{r}

cppFunction('NumericVector pdistC2(double x, NumericVector ys) {
              return sqrt(pow((x - ys), 2));
            }
')

x = 7.5
ys = rnorm(1e6)

microbenchmark(
  pdistR(x, ys),
  pdistC(x = x, ys = ys),
  sugar = pdistC2(x = x, ys = ys)
)

```

* Go over 'other useful functions' 

## STL

* go over stl in book

## Case Studies

* go over gibbs sampler (pg 424 in book)
* skip vectorization example

## Using Rcpp in a package

* go over in review
