---
title: "Functionals"
author: "Zach Schuster"
date: "8/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

### I will only be showing questions that may be insightful

* It's a mistake to focus on sped until you know it'll be a problem.

## Excercised 11.1.2

3. Use both for loops and lapply() to fit linear models to the mtcars using the formulas stored in this list:

```{r}
formulas <- list(
  mpg ~ disp,
  mpg ~ I(1 / disp),
  mpg ~ disp + wt,
  mpg ~ I(1 / disp) + wt
)

data(mtcars)

# for loop
for_models = vector(mode = 'list', length = length(formulas))
for (i in seq_along(formulas)) {
  for_models[[i]] = lm(formula = formulas[[i]], data = mtcars)
}

# lapply
lap_models = lapply(formulas, lm, data = mtcars)
```


## 11.2

* `sapply` is fine for interactive use, but a really bad idea for use in functions. use `vapply` whenever you can.

You may be more familiar with mapply() than Map(). I prefer Map() because:

* It’s equivalent to mapply with simplify = FALSE, which is almost always what you want.

* Instead of using an anonymous function to provide constant inputs, mapply has the MoreArgs argument that takes a list of extra arguments that will be supplied, as is, to each call. This breaks R’s usual lazy evaluation semantics, and is inconsistent with other functions.

In brief, mapply() adds more complication for little gain.

### Exercises 11.2.5

1. Use vapply() to:

* Compute the standard deviation of every column in a numeric data frame.

* Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)

```{r}

# numeric data
vapply(mtcars, sd, numeric(1))

# We will do the second piece of this question the way Hadley mentions and then
# also only using vapply once
data("iris")

# using 2 vapplys
vapply(iris[, vapply(iris, is.numeric, logical(1))],
       FUN = sd,
       numeric(1)
       )

# using 1 vapply (output isn't as clean but you also then see all columns)
vapply(iris,
       function(x) ifelse(is.numeric(x), sd(x), NA_real_),
       numeric(1))

```

2. Why is using sapply() to get the class() of each element in a data frame dangerous?

  * Class can return different length vectors which can break code


4. What does replicate() do? What sort of for loop does it eliminate? Why do its arguments differ from lapply() and friends?

`replicate is a wrapper for the common use of sapply for repeated evaluation of an expression (which will usually involve random number generation).`

The arguments differ because it is used to evaluate the same expression a number of times. Commonly used for random number generation


5. Implement a version of lapply() that supplies FUN with both the name and the value of each component.
```{r}
cm_lapply = function(X, FUN, nm, ...) {
  # browser()
  FUN <- match.fun(FUN)
    if (!is.vector(X) || is.object(X)) 
        X <- as.list(X)
    setNames(.Internal(lapply(X, FUN)), nm)
}

cm_lapply(letters[1:5], function(x) rnorm(1), nm = letters[1:5])
```


** Discuss Questions 6 and 7 out loud in group




