---
title: "Functionals"
author: "Zach Schuster"
date: "8/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

### I will only be showing questions that may be insightful

* It's a mistake to focus on speed until you know it'll be a problem.

## Excercised 11.1.2

3. Use both for loops and lapply() to fit linear models to the mtcars using the formulas stored in this list:

```{r}
formulas <- list(
  mpg ~ disp,
  mpg ~ I(1 / disp),
  mpg ~ disp + wt,
  mpg ~ I(1 / disp) + wt
)

data(mtcars)

# for loop
for_models = vector(mode = 'list', length = length(formulas))
for (i in seq_along(formulas)) {
  for_models[[i]] = lm(formula = formulas[[i]], data = mtcars)
}

# lapply
lap_models = lapply(formulas, lm, data = mtcars)
```


## 11.2

* `sapply` is fine for interactive use, but a really bad idea for use in functions. use `vapply` whenever you can.

You may be more familiar with mapply() than Map(). I prefer Map() because:

* It’s equivalent to mapply with simplify = FALSE, which is almost always what you want.

* Instead of using an anonymous function to provide constant inputs, mapply has the MoreArgs argument that takes a list of extra arguments that will be supplied, as is, to each call. This breaks R’s usual lazy evaluation semantics, and is inconsistent with other functions.

In brief, mapply() adds more complication for little gain.

### Exercises 11.2.5

1. Use vapply() to:

* Compute the standard deviation of every column in a numeric data frame.

* Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)

```{r}

# numeric data
vapply(mtcars, sd, numeric(1))

# We will do the second piece of this question the way Hadley mentions and then
# also only using vapply once
data("iris")

# using 2 vapplys
vapply(iris[, vapply(iris, is.numeric, logical(1))],
       FUN = sd,
       numeric(1)
       )

# using 1 vapply (output isn't as clean but you also then see all columns)
vapply(iris,
       function(x) ifelse(is.numeric(x), sd(x), NA_real_),
       numeric(1))

```

2. Why is using sapply() to get the class() of each element in a data frame dangerous?

  * Class can return different length vectors which can break code


4. What does replicate() do? What sort of for loop does it eliminate? Why do its arguments differ from lapply() and friends?

`replicate is a wrapper for the common use of sapply for repeated evaluation of an expression (which will usually involve random number generation).`

The arguments differ because it is used to evaluate the same expression a number of times. Commonly used for random number generation


5. Implement a version of lapply() that supplies FUN with both the name and the value of each component.
```{r}
cm_lapply = function(X, FUN, nm, ...) {
  # browser()
  FUN <- match.fun(FUN)
    if (!is.vector(X) || is.object(X)) 
        X <- as.list(X)
    setNames(.Internal(lapply(X, FUN)), nm)
}

x = cm_lapply(letters[1:5], function(x) rnorm(1), nm = letters[1:5])
```


** Discuss Questions 6 and 7 out loud in group

## Exercises 11.3.4

1. How does apply() arrange the output? Read the documentation and perform some experiments.

If each call to FUN returns a vector of length n, then apply returns an array of dimension c(n, dim(X)[MARGIN]) if n > 1. If n equals 1, apply returns a vector if MARGIN has length 1 and an array of dimension dim(X)[MARGIN] otherwise. If n is 0, the result has length 0 but not necessarily the ‘correct’ dimension.

If the calls to FUN return vectors of different lengths, apply returns a list of length prod(dim(X)[MARGIN]) with dim set to MARGIN if this has length greater than one.

In all cases the result is coerced by as.vector to one of the basic vector types before the dimensions are set, so that (for example) factor results will be coerced to a character array.

```{r}
(x = matrix(rnorm(12), nrow = 3))

apply(x, 1, sum)
apply(x, 2, sum)

apply(x, 1, function(x) x + 2) # tricky biz visually
# you could also get the above result as follows
t(x) + 2

apply(x, 2, function(x) x + 2)

```

2. There’s no equivalent to split() + vapply(). Should there be? When would it be useful? Implement one yourself.

By default vapply splits a list and applies a function to each split. I could see it being useful for a any sort of nested structure or a data frame that you'd like to apply a function by group. Below is a very naive approach. 

```{r}

svapply = function(x, FUN, FUN_VALUE, split_on) {
  vapply(split(x, split_on), FUN = FUN, FUN.VALUE = FUN_VALUE)  
}

data(mtcars)
svapply(mtcars$mpg, mean, numeric(1), mtcars$cyl)

```

3. Implement a pure R version of split(). (Hint: use unique() and subsetting.) Can you do it without a for loop?

This is implemented for vectors, not data.frames

```{r}
splitr = function(x, split_vec) {
  split_levels = sort(unique(split_vec))
  
  group_list = vector(mode = 'list', length = length(split_levels))
  
  for (i in seq_along(split_levels)) {
    group_list[[i]] = x[split_vec == split_levels[i]]
  }
  
  setNames(group_list, split_levels)
}

x = rnorm(100)
y = sample(LETTERS, 100, replace = TRUE)
str(splitr(x, y))

```


## Exercises 11.4.3

1. Why isn’t is.na() a predicate function? What base R function is closest to being a predicate version of is.na()?

`is.na` can return more than 1 value in certain scenarios and at times not return TRUE/FALSE. `anyNA` is much closer to a predicate for NA values. 

```{r}

is.na(NULL)
anyNA(NULL)

(tmp = data.frame(a = 1:2, b = c(NA, 4)))
is.na(tmp)
anyNA(tmp)
```

2. Use Filter() and vapply() to create a function that applies a summary statistic to every numeric column in a data frame.

```{r}
num_vapply = function(df, FUN, FUN_VALUE) {
  vapply(Filter(is.numeric, df), FUN = FUN, FUN.VALUE = FUN_VALUE)
}

data(iris)
num_vapply(iris, mean, numeric(1))
```

3. refer to online solutions

4. Implement Any(), a function that takes a list and a predicate function, and returns TRUE if the predicate function returns TRUE for any of the inputs. Implement All() similarly.

```{r}
# could also check to see if sum of vapply > 0 and not use `any`
Any = function(l, p) {
  any(
    vapply(l, p, logical(1))
  )
}

# could also check to see if sum of vapply == length(l) and not use `all`
All = function(l, p) {
  all(
    vapply(l, p, logical(1))
  )
}

# The above functions will generalize to a data.frame too
Any(iris, is.numeric)
All(iris, is.numeric)

```

5. Implement the span() function from Haskell: given a list x and a predicate function f, span returns the location of the longest sequential run of elements where the predicate is true. (Hint: you might find rle() helpful.)

```{r}

# test case will use custom function
test_func = function(x) all(x < 10)
test_list = unname(
  as.list(
    as.data.frame(replicate(1000, {sample(1:15, 3)}))
    )
  )

span = function(l, p) {
  if(!Any(l, p)) return(NULL)
  
  run_lengths = rle(vapply(l, p, logical(1)))
  pos = which(run_lengths$lengths == max(run_lengths$lengths[run_lengths$values == TRUE]) &
                    run_lengths$values == TRUE)
  # will only take the first location
  pos = pos[1]
  
  START = sum(run_lengths$lengths[1:(pos - 1)]) + 1
  STOP = START + run_lengths$lengths[pos] - 1
  len = run_lengths$lengths[pos]
  
  data.frame(start = START,
             stop = STOP,
             length = len)
}

span(test_list, test_func)
```


## Exercises 11.5.1

1. Implement arg_max(). It should take a function and a vector of inputs, and return the elements of the input where the function returns the highest value. For example, arg_max(-10:5, function(x) x ^ 2) should return -10. arg_max(-5:5, function(x) x ^ 2) should return c(-5, 5). Also implement the matching arg_min() function.

```{r}
arg_max = function(x, f) {
  
  f = match.fun(f)
  
  elements = f(x)
  x[which(elements == max(elements))]
}

arg_max(-10:5, function(x) x ^ 2)
arg_max(-5:5, function(x) x ^ 2)
```

*`arg_min` would look the exact same other than switching `max` to `min`

2. Will complete at a later time.
